---
title: "Homework 2"
author: "Jenesis Merriman"
date: "October 5, 2022"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Problem 1

Read and clean the data:

```{r read and clean}
transit_data = read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names()
```

Retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. Convert the entry variable from character (YES vs NO) to a logical variable (the ifelse or recode function may be useful):

```{r select and recode}
transit_data_selected = select(transit_data, line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada)

mutate(transit_data_selected, entry = recode(entry, "YES" = 1, "NO" = 0))
```


Write a short paragraph about this dataset – explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?

Answer the following questions using these data:

How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St 8th Avenue; 125st Broadway; 125st Lenox); the distinct function may be useful here.
How many stations are ADA compliant?
What proportion of station entrances / exits without vending allow entrance?
Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?

```{r}

```


```{r}

```

```{r}

```

