---
title: "Homework 2"
author: "Jenesis Merriman"
date: "October 5, 2022"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

The below code uses `readr` and `janitor::clean_names()` to import and clean the data. It also uses `select` to retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance and and `mutate` to convert the entry variable from character (YES vs NO) to a logical variable.

```{r read and clean}
transit_data = read_csv(
  file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
  col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>%
  janitor::clean_names() %>%
  select(line, station_name, station_latitude,
         station_longitude, route1:route11, entry, vending,
         entrance_type, ada)  %>%
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))
```

*This dataset focuses on NYC transit data and includes the variables line, station, name, station latitude, station longitude, routes served (1-11), entry, vending, entrance type, and ADA compliance. In order to clean the data, I imported the data, updated variable types, cleaned variable names, and selected the relevant columns. I then updated `entry` from `yes` / `no` to a logical variable. The resulting data set is 1868 rows x 19 columns. These data are not "tidy" in its current state. To further tidy our data, we would need to convert `route` variables from wide to long format.*

Answer the following questions using these data:

How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St 8th Avenue; 125st Broadway; 125st Lenox); the distinct function may be useful here.

```{r}

```

How many stations are ADA compliant?

```{r}

```

What proportion of station entrances / exits without vending allow entrance?

```{r}

```

Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?

```{r}

```


```{r}

```


```{r}

```

```{r}

```

