---
title: "Homework 2"
author: "Jenesis Merriman"
date: "October 5, 2022"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

The below code uses `readr` and `janitor::clean_names()` to import and clean the data. It also uses `select` to retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance and and `mutate` to convert the entry variable from character to a logical variable.

```{r}
transit_data = read_csv(
  file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
  col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>%
  janitor::clean_names() %>%
  select(line, station_name, station_latitude,
         station_longitude, route1:route11, entry, vending,
         entrance_type, ada)  %>%
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))
```

This dataset focuses on NYC transit data and includes the variables line, station, name, station latitude, station longitude, routes served, entry, vending, entrance type, and ADA compliance. In order to clean the data, I imported the data, updated variable types, cleaned variable names, and selected the relevant columns. I then updated `entry` from `yes` / `no` to a logical variable. The resulting data set is `r nrow(transit_data)` rows x `r ncol(transit_data)` columns. These data are not "tidy" in its current state. To further tidy our data, we would need to convert `route` variables from wide to long format.

**How many distinct stations are there?**

```{r}
transit_data %>% 
  select(station_name, line) %>% 
  distinct
```
*There are 465 distinct stations in the NYC transit system, according to our dataset.*

**How many stations are ADA compliant?**

```{r}
transit_data %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```
*There are 84 ADA compliant stations in the NYC transit system, according to our dataset.*

**What proportion of station entrances / exits without vending allow entrance?**

```{r}
transit_data %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```
*The proportion of station entrances / exits without vending that allow entrance is 0.377.*

**How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?** To answer this question, I will first reformat the data so that route number and route name are distinct variables:

```{r}
transit_data %>%
  pivot_longer(
    route1:route11,
    names_to = "route_number", 
    names_prefix = "route",
    values_to = "route_name") %>%
    filter(route_name == "A") %>%
  select(station_name, line) %>% 
  distinct

transit_data %>%
  pivot_longer(
    route1:route11,
    names_to = "route_number", 
    names_prefix = "route",
    values_to = "route_name") %>%
  filter(route_name == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```
*60 distinct stations serve the A train. Of these stations, 17 are ADA compliant.*

## Problem 2

Read and clean the Mr. Trash Wheel sheet, adding the variable 'sheet' to keep track of data for future combining:
```{r, warning=FALSE}
mr_trash = read_excel(
  "data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "Mr. Trash Wheel", range = 
    "A2:N535") %>% 
  janitor::clean_names() %>%
  mutate(sports_balls = as.integer(sports_balls), dumpster = as.double(dumpster), sheet =
           "mister") %>%
  drop_na()
```

Import, clean, and organize the data for Professor Trash Wheel, adding the variable 'sheet' to keep track of data for future combining::

```{r}
prof_trash = read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
    sheet = "Professor Trash Wheel", range = "A2:N117") %>% 
  janitor::clean_names() %>%
  mutate(sports_balls = as.integer(sports_balls), sheet = "professor") %>%
  drop_na()
```

Combine the Professor Trash Wheel dataset with the Mr. Trash Wheel dataset to produce a single tidy dataset:

```{r}
trash_tidy = full_join(mr_trash, prof_trash)
```

Write a paragraph about these data; you are encouraged to use inline R. Be sure to note the number of observations in the resulting dataset, and give examples of key variables. For available data, what was the total weight of trash collected by Professor Trash Wheel? What was the total number of sports balls collected by Mr. Trash Wheel in 2020?

## Problem 3

First, I will clean the data in pols-month.csv. This code will use `separate()` to break up the variable mon into integer variables year, month, and day; replace month number with month name; create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.

```{r}
pols = read_csv("data/fivethirtyeight_datasets/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(col = mon, into = c("year", "month", "day"), sep = "-") %>%
  select(-day) %>%
  mutate(
    month = month.name[as.numeric(month)], 
    year = as.numeric(year),
    prez_dem = recode(prez_dem, `1` = "dem", `0` = "gop"), 
    prez_gop = recode(prez_gop, `1` = "gop", `0` = "dem")) %>%
  pivot_longer(
    cols = starts_with("prez"),
    names_to = NULL,
    values_to = "president") %>%
  distinct()

print(pols)
```

Second, I will clean the data in snp.csv using a similar process to that above. For consistency across datasets, I will arrange according to year and month, and organize so that year and month are the leading columns.

```{r}
snp = read_csv("data/fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  separate(col = date, into = c("month", "day", "year"), sep = "/") %>%
  mutate(month = month.name[as.numeric(month)], year = ifelse(as.numeric(year) > 22,
                            as.numeric(year) + 1900, as.numeric(year) + 2000)) %>%
  select(year, month, close)

print(snp)
```

Third, I will tidy the unemployment data so that it can be merged with the previous datasets. This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.

```{r}
unemployment_tidy = read_csv("data/fivethirtyeight_datasets/unemployment.csv") %>%
  janitor::clean_names() %>%
  rename(
    January = jan,
    February = feb,
    March = mar,
    April = apr,
    May = may,
    June = jun,
    July = jul,
    August = aug,
    September = sep,
    October = oct,
    November = nov,
    December = dec) %>%
  pivot_longer(
    January:December,
    names_to = "month", 
    values_to = "unemployment_percentage")

print(unemployment_tidy)
```

Join the datasets by merging snp into pols, and merging unemployment into the result.

```{r}
pols_snp_merge = left_join(pols, snp)
print(pols_snp_merge)
```

```{r}
final_merge = left_join(pols_snp_merge, unemployment_tidy)
print(final_merge)
```

Write a short paragraph about these datasets. Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).
